[Back](lesson4_2.html)---
[to index](../../index.html)---
[Next](../lesson2/lesson5_1.html)

---

# Subsetting data in R

Subsetting data in R tends to be a tricky concept to grasp for
newcomers to R. However, understanding at least how to subset a data
frame is fundamental to understanding how to modify your data or select
just some of your larger dataset. We did some subsetting already in
the previous examples by calling a variable within a dataframe using
the 'Dollar sign' **\$**. Now we will discuss that and more.

First read in the data as we have been doing at the beginning of each lesson

```{r}

my_data <- read.csv("prewmort.csv")

```

Now let's select just one of the variables in the dataset and ask to
see it

```{r}

my_data$herdtype

```

You will see a sequence of zeros and ones in the console. What we did
was just ask to see the contents of the 'herdtype' variable in the
'my\_data' data frame. We returned one of the columns in the data
frame. We did this earlier as well when we tabulated and summarized
the contents of variables in the dataset, but we never really looked
as the sequence of numbers that we were passing to the functions to do
tabulation and summarizing. You could do the same thing for any of the
variables in the data frame.

Above we refer to the name of the variable in the data frame. We can
also refer to rows and columns in the data frame by their numerical
order called their 'index'. When you do this you no longer use the
**\$** operator but two square brackets, **[]**, after the name of the
data frame. Inside these brackets you put the indexes of the rows and
columns that you want back from the data frame separated by a comma.

```{r}

## mydata[row, column]
my_data[1,1]

```

We asked for the first row and first column in the data frame. Check
to make sure that makes sense by using the 'View' function to look at
your data and check the value of the first row, first column.

If you leave one of the two indexes blank R interprets this as wanted
all of the values. Therefore, if you were to do the following

```{r, eval = FALSE}

my_data[,]

```

R returns to you the entire data frame. If we use this principle, we
can ask for an entire row or entire column in the dataframe

```{r}

## the first row
options(width = 100)
my_data[1,]

## the first column
my_data[,1]

```

You can also supply a series of numbers, which in R is called a
'vector', to the index in order to retrieve several columns or rows at
a time. For example the first 3 rows:

```{r}

## generate a vector of 1 to 3
a <- c(1,2,3)

## This is what your vector looks like
a

## use this vector to index your dataframe
options(width = 100)
my_data[a,]

```

You should see the first 3 rows of your data frame.


In the above example we give the numeric index of the rows in the data
frame. We can also subset a dataset with a vector of TRUES and FALSES
that has the same length as the number of rows or columns in the
dataset.


All of this may seem abstract, but it is a very powerful foundation
for selecting subsections of you data for further analysis. Let's say
we want to examine only the data from breeding farms. We have the
variable 'herdtype' so we can use that to determine the indexes of the
rows of the data frame that we want.




## Now you try

This test indicated that we could not reject the null hypothesis that
herd type is independent of year. This was not a particularly
interesting test; try on you own to test whether preweaning mortality
is associated with parity. Have a look at the
[data description](../../about_datasets.html) if you need to know
which variables to look at.
