[Back](lesson4_1.html)---
[to index](../../index.html)---
[Next](../lesson2/lesson4_3.html)

---

# Tabulating data in R

Tabulating data is a useful way of looking at the frequencies of
categorical data and perhaps a cross-tabulation for investigating the
frequency of observations across multiple variables.

* Read in data from comma separated format to a dataframe named
  'my_data'.

```{r}

my_data <- read.csv("prewmort.csv")

```

* We will tabulate the variable 'herdtype' in the 'my_data' dataframe

```{r}

table(my_data$herdtype)

```

* In the previous lesson we used summarize to investigate the same
variable and in order to get an appropriate summary we needed to tell
R that the variable was a 'factor' (categorical). When the 'table'
function is used, R returns a table regardless of the data type
because that's what you asked for

* We can also ask R for a cross-tabulation by supplying a second
argument to the 'table' function.

```{r}

table(my_data$herdtype, my_data$year)

```

* This gives a table of the frequencies of each of the combinations of
  the variables.
* The table you just produced is 2x2; you may produce any size table:

A 1x14 table:

```{r}

table(my_data$sow_parity)

```

Or 14x20 table

```{r}

options(width = 100)
table(my_data$sow_parity, my_data$lsize)

```

Let's go back to the 2x2 table and quickly test do a chi-square test
on the table:

```{r}

my_table <- table(my_data$herdtype, my_data$year)

result <- chisq.test(my_table)

result

result$observed
result$expected
result$statistic
result$p.value

```

What we do above is first assign the result of the cross-tabulation to
a new object; saving the result in 'my_table' so we can use it for
something else. Then we submit this saved result , 'my_table', to the
'chisq.test' function and assign the output of that function to and
object: 'result'.

When we run the line: 'result', R returns a representation of the
content of the 'result' object. Here we can see the chi-squared
statistic and the corresponding _P_-value. We can also look closer at
the 'result' object and find that it contains the observed values that
we submitted to the 'chisq.test' function in 'result$observed'; the
expected values calculated by the function 'result$expected'; the
chi-squared statistic 'result$statistic' and the _P_-value in
'result$p.value'

This test indicated that we could not reject the null hypothesis that
herd type is independent of year. This was not a particularly
interesting test; try on you own to test whether
