---
output:
    html_document:
        css: ../../style.css
---

password: celestun

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 120)
```

# Practical task

In the current folder you have three datasets.

* farm_level.csv

* prewmort_a.csv

* prewmort_b.csv


Try some of the following tasks. You may wish to try to do this in an
Rmarkdown file in order to generate a report of your results or use an
R-script to do the analysis.

## Basic tasks

1. Explore the datasets. You will notice that two are at the
litter-level and one is at the farm-level. Join these datasets
together into one dataset.

2. Does your merging result in any missing values?

3. Are there farms that are in the farm_level dataset that are not in
the litter level dataset?

4. Investigate the effect of year on percent stillbirth in the litter. (model or graph)

5. Drop the observations from 1990

6. Generate a plot that can help you to understand the relationship
  between sow treatment and preweaning mortality. Is this relationship
  the same for different parities?



## Extra tasks:

1. Use the `plyr` package and the `ddply` function to count the number of observations within each level of a categorical variable. Try completing another operation on each piece of the collapsed dataframe.

The syntax is as follows:

```{r, comment = NA, eval =FALSE}
library(plyr)
ddply(.data, .var, .function)

```

2. In the third argument of the `ddply` function you gave a `function` as an argument. Part of the more advanced use of R is to organize your code into functions. Here is an example of a function that adds two numbers:

```{r, comment = NA}

my_function <- function(x, y){

    answer <- (x + y)
    return(answer)
    
    }

```

Evaluate the code above and try your new function:

```{r, comment = NA}

my_function(2,4)

```


a. Write a function that squares a single number, it will take a single argument
b. Write a function that runs multiple lines of code within the function and gives back the final result. Perhaps you would add and then multiply. 
