[Back](lesson1_3.html)---
[to index](../../index.html)---
[Next](../lesson3/lesson3_1.html)

---

# Getting your data into R

There are any number of format that you may use to store your data
that you want to summarize, plot or analyze using R. In this exercise
we will learn how to load data from STATA, excel, and from 2 different
text formats: comma separated and tab separated.

## Reading in data from comma separated format

When you read your data in from comma separated format then you have a
you just specify the new dataset name (my_data) and the file that you
want to read from. The default in the read.csv function is that the
first row in your .csv file contains the names of the variables.

```{r}

my_data <- read.csv("prewmort.csv")

## Have a look at the structure of the data with the 'str' function
str(my_data)

```

## Reading in data from tab delimited

When you read data from tab delimed form, you use a similar format to
.csv but now you must specify that the first row of your text file
contains the names of the variables by adding the argument 'header = TRUE'

```{r}

my_data <- read.table("prewmort.txt", header = TRUE)

## Have a look at the structure of the data with the 'str' function
str(my_data)

```


## Reading in data from Excel

Experience will teach you that storing your data in Excel format is
generally a bad idea since the Excel format changes over time and lots
of weird and wonderful formulae, colours and special characters can be
hidden from you. You may find that an apparently unaltered Excel file
does not contain the same formatting or data as a previous
version. You can avoid this problem by using a standardized text
format such as comma separated which can also be opened and edited in
Excel.

When you store your data in Excel, you have a spreadsheet or multiple
spreadsheets that are called worksheets in a workbook. When you import
data you need to specify the name of the file and index number of the
worksheet. In the example below we will load the first worksheet in
the 'prewmort.xlsx' workbook.

```{r}

## Use the readxl library
library(readxl)

my_data <- read_excel(path = "prewmort.xlsx", sheet = 1)

## Have a look at the structure of the data with the 'str' function
str(my_data)


```


View the data with the 'View' function. In R-studio you can also click
the name of the dataset in the environment tab in the top-right panel
of the editor


```{r, eval = FALSE}

View(my_data)
```

## Reading in data from STATA

Data that has been saved in STATA can be read into R using the
following function:

```{r}

## Use the foreign library
library(foreign)

my_data <- read.dta("prewmort.dta")

```

