# Explore and manage your data in R

This lesson will teach you

* Generating new variables
* Appending
* Merging

----

## Generate a new variable

```{r}
# Read data from file 'herd_20.txt'
herd_20 <- read.table("herd_20.txt", header = TRUE)

# View a summary of the dataset
summary(herd_20)

# View the first six rows of the dataset
head(herd_20)

# View the structure of the dataset
str(herd_20)
```

Lets add a column with the name of the herd manager

```{r}
herd_20$manager <- "Alice"

# View the structure of the dataset again
str(herd_20)

# View the first six rows of the dataset
head(herd_20)
```

## Append data to a data.frame

To combine the 16 data files in this lesson, we will introduce the
[for-loop](https://cran.r-project.org/doc/manuals/r-release/R-intro.html#Loops-and-conditional-execution). Using
the `for`, we can read one file at a time and append to the result. If
you browse the folder for this lesson, you should find the following
16 files.

```{r}
files <- c("herd_20.txt", "herd_21.txt", "herd_22.txt", "herd_23.txt",
           "herd_24.txt", "herd_25.txt", "herd_26.txt", "herd_27.txt",
           "herd_28.txt", "herd_30.txt", "herd_31.txt", "herd_33.txt",
           "herd_34.txt", "herd_36.txt", "herd_37.txt", "herd_38.txt")
```

Lets start to explore the `for` loop before reading data.

```{r}
for(i in 1:10) {
    cat("Hello R! ")
}
```

## Merge data to a data.frame

